---
title: 'Multiverse Analysis with `specr`'
subtitle: 'Exercise II'
author: "Philipp Masur"
date: "2023-10"
output: 
  github_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r, echo=F, message=F}
knitr::opts_chunk$set(echo = TRUE, results = FALSE, message = FALSE, warning = FALSE, fig.keep = 'none')
library(printr)
```


# Introduction

This tutorial shows how to use the package `specr` to run specification curve analyses (SCA). The basic idea of SCA is to solve the problem that there always exist a large number of researcher's degrees of freedom when analyzing a data set. Data are actively "shaped" during the transformation from raw data to analyzable form and researchers often apply multiple processing steps during data preparation. These steps involve numerous decisions, offering various, alternative options at each stage. Taking all of these (often arbitrary) decisions into account, it is important to acknowledge that they generate multiple versions based on the choices made — a data multiverse! The problem is that each dataset in this multiverse can produce different statistical outcomes - a statistical multiverse. To solve this problem and provide a systematic assessment of the resulting "garden of forking paths", SCA usually consists of 2 to 3 steps:


1. Define the set of reasonable specifications to estimate
2. Estimate all specifications and visualize/analyze results
3. Conduct joint statistical tests using an inferential specification curve

`specr` provides easy to use function to conduct all three steps (step three only in the development version). 

# Preparation

For this tutorial, we are going to load the package collection `tidyverse` (for data wrangling and visualization) and the package `specr`.

```{r}
library(tidyverse)
# install.packages("specr")
library(specr)
library(psych)
```


## Getting data

To have some data to play around with, we are going to use data from an experiment by Masur, DiFranzo and Bazarova (2021) in which they studied the influence of the collective norm on subsequent social norm perceptions, and self-disclosure behavior on social media. More specifically, they exposed participants to social media feeds that varied with regard to the prevalence of visual disclosure in posts and profiles (3 x 3 between-subject experiment). Subsequently, participants reported descriptive, injunctive, and subjective norm perceptions and the likelihood with which they would disclose themselves visually in such a network. 

**Reference:**

- Paper: Masur, P. K., DiFranzo, D. J., & Bazarova, N. N. (2021). Behavioral Contagion on Social Media: Effects of Social Norms, Design Interventions, and Critical Media Literacy on Self-Disclosure. PLOS One, 16(7). e0254670. https://doi.org/10.1371/journal.pone.0254670
- Data and Code: https://osf.io/qxjsp/


```{r}
d <- read.csv("https://osf.io/kfm6x/download", 
              header = TRUE, na.strings = "NA") %>% 
  as_tibble

head(d)
```

## Understanding the data set

The data set consists of the following variables:

- **condition**: What condition the participant was assigned to
- **norm**: First manipulation, amount of posts (in %) that included visual disclosure in the feed
- **profile**: second manipulation, amount of profile pictures (in %) that included visual disclosure
- **MLI_**: Media literacy (20 items)
- **CRI_**: Critical thinking disposition (9 items)
- **ESL_NOR_**: Norm perceptions (1-4: descriptive, 5-8: injunctive, 9-1: subjective)
- **ESL_BEH_**: Self-disclosure intention
- **Socio-Demographics**: age, gender, education,...

```{r}
# Sample description
describe(d$age)
table(d$gender) # 0 = male, 1 = female, 2 = other

# Experimental conditions
d %>%
  group_by(norm, profile) %>% 
  tally                       

# Variables
names(d)
```


## Creating variables

For the purpose of this tutorial, we are going to create several mean indices that represent the concepts of interest. For the dependent variable (i.e. self-disclosure intentions), we are quickly going to check if the scale can be improved by deleting one or two items. 

```{r}
alpha(d %>% select(contains("ESL_BEH")))
```

According to the reliability analysis, we could exclude item 06 to improve the internal consistency (alpha). We hence compute both the full and a reduce scale. 

```{r}
# Preparations
d <- d %>%
  mutate(descriptive_norm = rowMeans(d %>% select(ESL_NOR_01:ESL_NOR_04)),
         injunctive_norm = rowMeans(d %>% select(ESL_NOR_05:ESL_NOR_08)),
         subjective_norm = rowMeans(d %>% select(ESL_NOR_09:ESL_NOR_12)),
         social_norm = rowMeans(d %>% select(contains("ESL_NOR"))),
         self_disclosure_all = rowMeans(d %>% select(ESL_BEH_01:ESL_BEH_06)),
         self_disclosure_red = rowMeans(d %>% select(ESL_BEH_01:ESL_BEH_05)),
         media_literacy = rowMeans(d %>% select(contains("MLI_CRI"))),
         critical_thinking = rowMeans(d %>% select(contains("CRI_THI"))),
         gender = ifelse(gender > 1, NA, gender),
         gender = factor(gender, labels = c("male", "female")))
```

# Specification curve analysis

## Setting up specifications

The first step of any specification curve analysis is to set up the specifications. In `specr`, we do this using the function `setup()`. There a endless way to specify different decisions (see this [vignette](https://masurp.github.io/specr/articles/different-specifications.html) for a more detailed tutorial). In this case, we are going to vary the type of norm perception (as independent variable), the scale measuring self-disclosure (as dependent variable), the covariates to include, and the subset analyses that we want to include. We can check the specifications using the generic `summary()` function. 

```{r}
specs <- setup(data = d,
               x = c("descriptive_norm", "injunctive_norm", "subjective_norm", "social_norm"),
               y = c("self_disclosure_all", "self_disclosure_red"),
               model = "lm",
               controls = c("media_literacy", "critical_thinking", "age"),
               subsets = list(gender = c("male", "female")))
summary(specs)
```

The output contains quite some information about our specification setup such as the number of specification (192), the decisions involved, the function used to extract parameters (could be changed) and also shows the first 6 specifications in the specification table. 

## Running the analysis

To estimate all models, we simply use the function `specr()` around this specification object. Depending on the complexity of the estimations and the size of the data set, this can take a while (parallization is possible, see this [vignette](https://masurp.github.io/specr/articles/parallelization.html)). We again can explore the results using the generic `summary()` function. 

```{r}
results <- specr(specs)
summary(results)
```

We can see that across all 192 specifications, the resulting median coefficient for the relationship between social norm perceptions and self-disclosure intention is Mdn = 0.54 (MAD = 0.06). 
We further see that sample sizes ranged from 254 to 670 (bear in mind that the specifications include subset analyses). 

## Visualizing the curve

Most commonly, specification curve analysis are reported using a visualization. We can obtain this visualization (a combination of the specification curve and the choice panel) with the generic `plot()` function. 

```{r}
plot(results)
```

We can customize the appearance of this plot in many ways:

```{r}
a <- plot(results, type = "curve", ribbon = T, ci = F)
b <- plot(results, type = "choices", choices = c("x", "y", "controls", "subsets"))
c <- plot(results, type = "samplesizes")

# Combine plots
plot_grid(a, b, c,
          nrow = 3, axis = "tblr",
          align = "hv", 
          rel_heights = c(1, 2, .75))
```


**Exercise:** Can you plot a similar plot that looks at the distributions of r-square instead of the estimate? Use the help function for `?plot.specr.object` if necessary.

```{r}
# Code here
```

## Alternative visualizations

Despite the prominence of this visualization in the literature, we can visualize our results in various other ways. After all, the result is just a distribution of effect sizes (and their 95% confidence intervals). 

```{r}
# Box plot visualization
plot(results, type = "boxplot")

# Density distributions
library(ggridges)
results %>%
  as_tibble %>%
  ggplot(aes(x = estimate, y = x)) +
  geom_density_ridges2(aes(fill = y), alpha = .5, scale = .6) +
  theme_minimal()
```

## Summarization techniques

Descriptively, we can also summarize the effect size distribution using the generic `summary()` function. 

```{r}
summary(results, type = "curve")
summary(results, type = "curve", group = "x")
summary(results, type = "curve", group = c("x", "y"))
summary(results, type = "curve", group = "controls",
        stats = list(mean = mean, 
                     median = median))
```


# Inference for specification curve analysis

If we are convinced that our specifications represent a meaningful sample of the entire set of reasonable, non-redundant specifications, we can also run inference tests on the specification curve. To do this, we need to install the development version of `specr`. 

## Rerunning the model with a custom function

First, we have to rerun the specification curve with a custom extract function to make sure that we get the entire model object from each specification (and not just the parameter of interest). 

```{r}
# Install development version
# devtools::install_github("masurp/specr")  
# library(specr)

# Custom function
tidy_new <- function(x) {
  fit <- broom::tidy(x, conf.int = TRUE)
  fit$res <- list(x)  # Store model object
  return(fit)
}

# Setup with function
specs <- setup(data = d,
               x = c("descriptive_norm", "injunctive_norm", "subjective_norm", "social_norm"),
               y = c("self_disclosure_all", "self_disclosure_red"),
               model = "lm",
               controls = c("media_literacy", "critical_thinking", "age"),
               fun1 = tidy_new)
# Check
summary(specs)

# Estimate models
results <- specr(specs)

# Check
summary(results)
```

## Resampling under-the-null

Generating distributions for these test statistics under the null hypothesis isn’t feasible analytically, but we can create these distributions through resampling under the null hypothesis. This process entails adjusting the observed data to ensure a known true null hypothesis and then randomly sampling from the modified data. The test statistic is then computed on each of these samples. 

We can secondly use the function `boot_null` to compute "under-the-null" data sets and re-compute SCA results on them. This process is akin to bootstrapping and thus takes a while, particularly if the number of samples is high. We run only 10 samples for this example, but in practice, we should rather use n = 500 or even n = 1,000.

```{r, eval = F}
# To make it reproducible
set.seed(42)

# Run bootstrapping (takes a couple of minutes)
boot_models <- boot_null(results, specs, n_samples = 3) 

boot_models$boot_models %>%
  unnest(cols = c("fit")) %>%
  select(id, x, y, controls, estimate) %>%
  print(n = 70)
```

## Inference tests

Lastly, if we use the `summary()` function on the result `specr.boot` object, we get the desired inference tests:

1. A test for whether the median effect across all specifications is more extreme than would be expected if all specifications had a true effect of zero
2. A test whether the share of specifications with statistically significant effects in the expected direction is more extreme than would be expected if all specifications had an effect of zero

```{r, eval = F}
summary(boot_models)
```

As we can see, the data is more extreme that would be expected if all specifications had an effect of zero. 

## Visualizing the inference curve

As shown in Simonsohn et al., (2020), we can also visualize the distribution of the curve under-the-null and compare it with the actually observed specification curve. We can simply do this by plotting the `specr.boot` object.

```{r, eval = F}
plot(boot_models)
```

In this case, the observed curve is clearly outside the resampled curves under-the-null.  


**Exercise:** Create a full specification curve analysis for the relationship between media literacy and self-disclosure. With regard to potential specifications...

- investigate whether the 20 items actually form a uni-dimensional model (you will find that it does not) and include several dimensions, measures with different items, etc.
- include a variety of covariates (most imporantly the manipulations: norm and profile and their interaction...)
- create subset analyses (e.g., age groups, gender, educational levels...)
- are there different model estimation approaches?
- What about outliers?
- What about imputation?


```{r}
# Code here
```


